\textbf{\sc Deliberative Autonomy for Marine Robotics}

### 2012-01-20 trunk/europa.tex
added more latex markup to europa.tex and included europa.tex in setobook.tex

	- Bounds propagation: To infer upper and lower variable bounds. For example, from the constraints x1 + x2 <= 2, xi>=0, we can infer [0,2] bounds for x1 and x2

	- Domain reduction: To infer a valid set of values for a variable. For example, for discrete variables, the constraints all_different(x1,x2,x3) and xi>=0, xi <=4. if x1 =1 and x2=3 we can infer that the valid domain for x3 is reduced to {2,4}

@@

- The Constraint Reasoning Engine (CRE): manages variables, the domains from which they can take values, and constraints that define relationships among variables. It also provides an efficient arc consistency mechanism (TODO: ref to AC-3). The CRE is designed so that specialized reasoning algorithms for specific constraints can be easily and efficiently plugged in.

- The Plan Database (PDB): manages object types and objects, which are a mechanism to group variables to more naturally model the real world in much the same way that object oriented programming does. Also manages tokens, which are a mechanism to group variables to represent temporally scoped state. Objects and tokens can be used to model planning domains in a much more natural and extensible way that a pure CSP approach can (TODO: provide snippet?)

- The Rules Engine (RE): manages causal dependencies in a planning model (TODO: explain)

- The language modules: NDDL, ANML, and other higher level modeling languages can me used to create domain models and problem instances (TODO: point advantages over pure programming API approaches, provide snippets)

- API Layer: makes all the services available so that EUROPA can be used to build applications.

- Extension Modules bundled with EUROPA: Temporal Network, Resources, NDDL, ANML.

@@

- A domain model that describes the variable, object, predicate and action types that are relevant for the problem.
- A problem instance (also called initial state) that consists of:
	- Variable and Object instances that exist for the entire planning horizon.
	- Temporally scoped predicate and action instances.
	- Temporally scoped goals.

@@

EUROPA provides a built-in solver that performs Plan Space Planning (TODO: ref), in this approach, the initial state is considered a partial plan that needs to be refined toward a solution plan that achieves the goals. The operations to refine the partial plan PP at any time are:
	1- Find the flaws of PP, that is the conditions that prevent it from being a solution plan (flaw types are explained below).
	2- Select one such flaw
	3- Select a resolver for the flaw
	4- Refine PP by applying the resolver
	5- If an inconsistency is found, try another resolver.
	6- If all possible resolvers for a particular flaw fail, return failure, otherwise continue until resolving all the flaws.

@@

	1. Unbound Variable: a variable in the partial plan whose domain is not a singleton. Unbound Variables are resolved by specification of a value from the domain of the variable.
	2. Open Condition: an open condition is an inactive token. Inactive tokens can be generated by explicitly posted goals, or when a token is activated, rules may fire that create inactive slave tokens (TODO: point to more detailed decription above in Rules Engine section).  Open Conditions can be resolved in three ways: 
		- Merging: A free activity is merged with a matching activity already in the plan. An activity a is said to match an activity a? if a and a? unify and the temporal constraints involving a are satisﬁed by a?. Thus,  a and a? can be considered the same activity and we do not need to introduce a in the plan. Consequently, the compatibilities associated with a are not ﬁred, because they have been already triggered when a? was introduced in the plan. 
		- Activation We introduce a new activity a in the current plan associating it with the proper timeline, but without choosing a speciﬁc time slot for it.The compatibilities associated with a are applied and the subgoal activities resulting from those compatibilities are introduced as free activities. This results in both an ordering ﬂaw, corresponding to the just activated activity, and a number of open condition ﬂaws, corresponding to the new subgoal activities. 
		- Rejection: Some goals may be optional, if that is the case for a particular order condition, its corresponding flaw can be resolved by discarding the inactive token.

	3. Threat: once a token has been placed in the partial plan it may impact other tokens indirectly through possible overlapping requirements on objects. Recall for example that a token may belong to objects (e.g. Timelines) which require a total order over their tokens. If any 2 tokens could possibly overlap (though not necessarily), then they pose a threat to each other in terms of achieving an extension of the current partial plan which is complete and consistent. Similarly, threats may arise where tokens share a common resource and their current state might yield extensions of the current partial plan which are inconsistent. Threats are resolved by imposing ordering constraints among tokens.

@@

// Function to solve a partial plan by successive refinement
// until all flaws are resolved.
bool solve(PartialPlan plan)
{
	// Propagate the constraints to test for inconsistency. If found to be
	// inconsistent, then we can return false since this is a dead-end i.e. no refinements
	// to p can yield a consistent plan.

	if(isInconsistent(plan))
		return false;

	// Non-deterministically choose a flaw from the set of available flaws.
	Flaw flaw = chooseFlaw(plan);

	// If there are no flaws, then p is complete and we can terminate with success.
	if(flaw == NULL)
		return true;

	// Otherwise we formulate a decision point which is a branch in the search space. Each
	// choice is a particular refinement operation and the DecisionPoint collects all possible
	// refinement operations for the given flaw.
	DecisionPoint decision = makeDecisionPoint(flaw, plan);

	// Continue as long as we have something to try
	while(decision.hasNext()) {
		// A new partial plan is obtained by application of a refinement operator. Note that the ordering
    		// over refinement operators to select is a non-deterministic step.
		PartialPlan pp = decision.executeNext();

		// Recursive call to solve the new planning problem. If successful, then we are done.
		if(solve(pp))
			return true;
		else // Otherwise, retract the last refinement operation
			d.undo();
  	}

	// If we arrive here, then we have exhausted all options to resolve the flaw, including the case where
  	// no options were available initially. Thus the problem cannot be solved.
  	return false;
}

@@

The algorithm takes as input a partial plan p and returns true if a complete and consistent refinement of p could be found (or if p is initially complete and consistent), and false otherwise. This algorithm provides for a sound and complete search, assuming that no flaws or available refinement operators are pruned unnecessarily. Dead-ends in the search are discovered through constraint propagation. Constraint propagation is a vehicle for evaluating the consistency of a partial plan and also for filtering infeasible values from consideration prior to commitment, allowing in some cases a strong look-ahead capability which is essential for tractable search. Consistency testing is initiated by the isConsistent procedure. The algorithm permits a heuristically controlled search by applying orderings for chooseFlaw and makeDecisionPoint. It results in a chronologically-backtracking, search.

>>>>

\begin{enumerate}
	\item Bounds propagation: To infer upper and lower variable bounds. For example, from the constraints x1 + x2 <= 2, xi>=0, we can infer [0,2] bounds for x1 and x2

	\item Domain reduction: To infer a valid set of values for a variable. For example, for discrete variables, the constraints allDifferent(x1,x2,x3) and xi>=0, xi <=4. if x1 =1 and x2=3 we can infer that the valid domain for x3 is reduced to {2,4}

\end{enumerate}

@@

\begin{enumerate}

	\item The Constraint Reasoning Engine (CRE): manages variables, the domains from which they can take values, and constraints that define relationships among variables. It also provides an efficient arc consistency mechanism (TODO: ref to AC-3). The CRE is designed so that specialized reasoning algorithms for specific constraints can be easily and efficiently plugged in.

	\item The Plan Database (PDB): manages object types and objects, which are a mechanism to group variables to more naturally model the real world in much the same way that object oriented programming does. Also manages tokens, which are a mechanism to group variables to represent temporally scoped state. Objects and tokens can be used to model planning domains in a much more natural and extensible way that a pure CSP approach can (TODO: provide snippet?)

	\item The Rules Engine (RE): manages causal dependencies in a planning model (TODO: explain)

	\item The language modules: NDDL, ANML, and other higher level modeling languages can me used to create domain models and problem instances (TODO: point advantages over pure programming API approaches, provide snippets)

	\item API Layer: makes all the services available so that EUROPA can be used to build applications.

	\item Extension Modules bundled with EUROPA: Temporal Network, Resources, NDDL, ANML.

\end{enumerate}

@@

\begin{enumerate}

\item A domain model that describes the variable, object, predicate and action types that are relevant for the problem.
\item A problem instance (also called initial state) that consists of:
	\item Variable and Object instances that exist for the entire planning horizon.
	\item Temporally scoped predicate and action instances.
	\item Temporally scoped goals.

\end{enumerate}

@@

EUROPA provides a built-in solver that performs Plan Space Planning (TODO: ref), in this approach, the initial state is considered a partial plan that needs to be refined toward a solution plan that achieves the goals. The operations to refine the partial plan PP at any time are:
\begin{enumerate}
	\item Find the flaws of PP, that is the conditions that prevent it from being a solution plan (flaw types are explained below).
	\item Select one such flaw
	\item Select a resolver for the flaw
	\item Refine PP by applying the resolver
	\item If an inconsistency is found, try another resolver.
	\item If all possible resolvers for a particular flaw fail, return failure, otherwise continue until resolving all the flaws.
\end{enumerate}

@@

\begin{enumerate}
	\item Unbound Variable: a variable in the partial plan whose domain is not a singleton. Unbound Variables are resolved by specification of a value from the domain of the variable.

	\item Open Condition: an open condition is an inactive token. Inactive tokens can be generated by explicitly posted goals, or when a token is activated, rules may fire that create inactive slave tokens (TODO: point to more detailed decription above in Rules Engine section).  Open Conditions can be resolved in three ways: 
\begin{enumerate}
		\item Merging: A free activity is merged with a matching activity already in the plan. An activity a is said to match an activity a? if a and a? unify and the temporal constraints involving a are satisﬁed by a?. Thus,  a and a? can be considered the same activity and we do not need to introduce a in the plan. Consequently, the compatibilities associated with a are not ﬁred, because they have been already triggered when a? was introduced in the plan. 
		\item Activation: We introduce a new activity a in the current plan associating it with the proper timeline, but without choosing a speciﬁc time slot for it.The compatibilities associated with a are applied and the subgoal activities resulting from those compatibilities are introduced as free activities. This results in both an ordering ﬂaw, corresponding to the just activated activity, and a number of open condition ﬂaws, corresponding to the new subgoal activities. 
		\item Rejection: Some goals may be optional, if that is the case for a particular order condition, its corresponding flaw can be resolved by discarding the inactive token.
\end{enumerate}

	\item Threat: once a token has been placed in the partial plan it may impact other tokens indirectly through possible overlapping requirements on objects. Recall for example that a token may belong to objects (e.g. Timelines) which require a total order over their tokens. If any 2 tokens could possibly overlap (though not necessarily), then they pose a threat to each other in terms of achieving an extension of the current partial plan which is complete and consistent. Similarly, threats may arise where tokens share a common resource and their current state might yield extensions of the current partial plan which are inconsistent. Threats are resolved by imposing ordering constraints among tokens.

\end{enumerate}

@@

01.  bool solve(PartialPlan plan)
02.   {
03.      if (isInconsistent(plan))  return false;
04.
05.      Flaw flaw = chooseFlaw(plan);
06.      if (flaw == NULL) return true;
07.
08.      DecisionPoint decision = makeDecisionPoint(flaw, plan);
09.
10.      while(decision.hasNext()) {
11.          PartialPlan pp = decision.executeNext();
12.
13.          if (solve(pp)) return true;
14.          else d.undo();
25.      }
16.
17.      return false;
18.  }

@@

The algorithm takes as input a partial plan p and returns true if a complete and consistent refinement of p could be found (or if p is initially complete and consistent), and false otherwise.

\begin{enumerate}
	\item Line 3: Propagate the constraints to test for inconsistency. If found to be inconsistent, then we can return false since this is a dead-end i.e. no refinements to p can yield a consistent plan.
	\item Line 5: Choose a flaw from the set of available flaws.
	\item Line 6: If there are no flaws, then p is complete and we can terminate with success.
	\item Line 8: Otherwise we formulate a decision point which is a branch in the search space. Each choice is a particular refinement operation and the DecisionPoint collects all possible refinement operations for the given flaw.
	\item Line 10: We keep trying until the chosen flaw is resolved, or until we run out of resolvers to try.
	\item Line 11: A new partial plan is obtained by application of a refinement operator. Note that the ordering over refinement operators to select is a non-deterministic step.
	\item Line 13: Recursive call to solve the new planning problem. If successful, then we are done.
	\item Line 14: Otherwise, retract the last refinement operation and move on to try the next one.
	\item Line 17:  If we arrive here, then we have exhausted all options to resolve the flaw, including the case where no options were available initially. Thus the problem cannot be solved.
\end{enumerate}

This algorithm provides for a sound and complete search, assuming that no flaws or available refinement operators are pruned unnecessarily. Dead-ends in the search are discovered through constraint propagation. Constraint propagation is a vehicle for evaluating the consistency of a partial plan and also for filtering infeasible values from consideration prior to commitment, allowing in some cases a strong look-ahead capability which is essential for tractable search. Consistency testing is initiated by the isConsistent procedure. The algorithm permits a heuristically controlled search by applying orderings for chooseFlaw and makeDecisionPoint. It results in a chronologically-backtracking, search.
